---
title: "R Machine Learning Assignment"
output: html_document
---

# Summary
We are going to build a machine learning model to predict 

# Load Data
```{r}
training <- read.csv("data/pml-training.csv", na.strings=c("NA", "#DIV/0!"), stringsAsFactors=FALSE)
training$classe <- as.factor(training$classe)
to_predict <- read.csv("data/pml-testing.csv", na.strings=c("NA", "#DIV/0!"), stringsAsFactors=FALSE)
```
We see that data has 159 predictor variables. The variable to be predicted is "classe". We see that a lot of predictor variables are NA for some rows. Instead of averaging, we will choose to remove any columns with NA values.
```{r}
training_cc <- training[, colSums(is.na(training)) == 0]
to_predict <- to_predict[, which(names(to_predict) %in% names(training_cc))]
```
Here we have filtered out predictors for which there are no NA values in the testing set. Still we can see that we have `r dim(training_cc)[2]` predictor variables left now. We will now partition this data into our own training set and testing set for cross-validation. We are ignoring first six columns in the training set because they are not going to be used in our model.
```{r}
library(caret)
library(randomForest)
set.seed(156)
inTrain <- createDataPartition(y=training_cc$classe, p=0.6, list=FALSE)
my_training <- training_cc[inTrain, 7:60]
my_testing <- training_cc[-inTrain, 7:60]
model <- randomForest(classe ~ ., data=my_training, importance=TRUE, proximity=TRUE)
model
```
Now that we have a random forest model, we can try to estimate in-sample and out of sample errors.
In sample error is nothing but the accuracy on the training set. 
```{r}
sum(model$predicted == my_training$classe)/dim(my_training)[1]
```
Out of sample error can be estimated using the training set we created when we partitioned the original training set.
```{r}
sum(predict(model, my_testing) == my_testing$classe)/dim(my_testing)[1]
```
We will use our model to predict the held out testing set.
```{r,eval=FALSE}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
answers <- predict(model, to_predict)
pml_write_files(answers)
```

## Cross Validation
We use k-fold cross-validation to estimate out of sample error. 


